% Springer LNCS — Consent to Publish (recreation for ICAISC 2026)
% Print, sign by hand, scan to PDF, pack as 12054_cp.zip
\documentclass[11pt]{article}
\usepackage[a4paper, margin=2.5cm]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{enumitem}
\usepackage{array}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{xcolor}

\pagestyle{empty}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt}

\begin{document}

\begin{center}
{\Large\bfseries Consent to Publish}\\[4pt]
{\large Springer-Verlag GmbH — Lecture Notes in Computer Science}\\[12pt]
\end{center}

\hrule
\vspace{8pt}

\textbf{Title of the Book/Volume:}\\
Artificial Intelligence and Soft Computing --- ICAISC 2026\\
Lecture Notes in Artificial Intelligence (LNAI), Springer

\vspace{6pt}
\textbf{Volume Editor(s):} Leszek Rutkowski et al.

\vspace{6pt}
\textbf{Title of the Contribution:}\\
Hybrid Neighborhood Evaluation in Tabu Search and Simulated Annealing for the Permutation Flow Shop Problem

\vspace{6pt}
\textbf{Author(s) of the Contribution:}\\
Remigiusz Wojew\'{o}dzki, Wojciech Bo\.{z}ejko

\vspace{6pt}
\textbf{Corresponding Author:}\\
Remigiusz Wojew\'{o}dzki\\
Wroc{\l}aw University of Science and Technology, Poland\\
\texttt{remigiusz.wojewodzki@gmail.com}

\vspace{12pt}
\hrule
\vspace{12pt}

When Author is more than one person the expression ``Author'' as used in this
agreement shall apply collectively unless otherwise indicated.

\vspace{6pt}
\textbf{§\,1 Rights Granted.}
The Author hereby grants to Springer-Verlag GmbH, Heidelberger Platz~3,
14197~Berlin, Germany (hereinafter called ``Springer'') the exclusive,
sole, permanent, world-wide, transferable, sub-licensable and unlimited right
to reproduce, publish, distribute, transmit, make available or otherwise
communicate to the public, translate, publicly perform, archive, store,
lease or lend and sell the Contribution or parts thereof individually or
together with other works in any language, in all revisions and versions
(including soft cover, book club and collected editions, anthologies,
advance printing, reprints or print-on-demand, microfilm editions,
audiograms and videograms), in all forms and media of expression
including in electronic form (including offline and online use, push or
pull technologies, use in databases and networks for display, print and
storing on any and all stationary or portable end-user devices, e.g.\
e-books, e-journals, e-papers, e-archive, and the like), in all other
forms and media of expression now known or developed in the future,
including the right to produce, or have produced, the Contribution or
parts thereof in all customary formats for promotional purposes, for the
Contribution identified above.

\vspace{6pt}
\textbf{§\,2 Rights Retained by Author.}
The Author retains, in addition to uses permitted by law, the right to
communicate the content of the Contribution to other scientists, to share
the Contribution with them in manuscript form, to perform or present the
Contribution or to use the content for non-commercial internal and
educational purposes, provided the Springer publication is mentioned as the
original source of publication in any printed or electronic materials.

The Author may self-archive an Author's accepted manuscript of their
Contribution on their own personal website and/or in their institutional
repository, and on the preprint server arXiv or a similar server,
provided the acknowledgement to the original source of publication is
given.

\vspace{6pt}
\textbf{§\,3 Warranties.}
The Author warrants that the Contribution is original and that the Author
has full power to make this agreement. The Author signs for and accepts
responsibility for releasing this material on behalf of any and all
co-authors. The Author warrants that the Contribution contains no
libelous or unlawful statements, does not infringe upon the rights of
others, and that the Author has obtained written permission for the use
of any material from other sources as necessary.

\vspace{6pt}
\textbf{§\,4 Third-Party Material.}
If the Contribution includes third-party material, the Author has
obtained permission from the copyright holder to use such material.

\vspace{16pt}
\hrule
\vspace{16pt}

\textbf{Signature of the Corresponding Author:}

\vspace{30pt}

\begin{tabular}{@{}p{0.45\textwidth}p{0.45\textwidth}@{}}
\rule{6cm}{0.4pt} & \rule{6cm}{0.4pt} \\
Name (print): Remigiusz Wojew\'{o}dzki & Date: \hspace{3cm} \\[20pt]
\rule{6cm}{0.4pt} & \\
Signature & \\
\end{tabular}

\vspace{20pt}

\textbf{Co-Author signature (if applicable):}

\vspace{30pt}

\begin{tabular}{@{}p{0.45\textwidth}p{0.45\textwidth}@{}}
\rule{6cm}{0.4pt} & \rule{6cm}{0.4pt} \\
Name (print): Wojciech Bo\.{z}ejko & Date: \hspace{3cm} \\[20pt]
\rule{6cm}{0.4pt} & \\
Signature & \\
\end{tabular}

\vspace{20pt}
{\small\textit{Paper ID: 12054}}

\end{document}
